BivData Setup

### 2020-05-31 src/BivData_Setup.Rmd
more updates to figure scripts ect

#BIVDATA MASTER SHEET
new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")
#83975
#83976
#83009
#80898
#80919
#61084
#60241
#60830

#Clean away all the empty rows
new.Curate.FULL <- new.Curate.FULL[!(is.na(new.Curate.FULL$fileName) | new.Curate.FULL$fileName==""), ]
#this takes a while to run

new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file
#make OG copy
new.Curate.FULL.org <- new.Curate.FULL
new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

@@

#SC PASS
new.Curate.FULL <- new.Curate.FULL[new.Curate.FULL$SC.pass == '1',]#
#10480 #9819, #9769

#fix these duplicated ones
#remove duplicates
Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151

Curated_BivData <- add_mouse(Curated_BivData)
Curated_BivData <- add_category(Curated_BivData)
Curated_BivData <- add_strain(Curated_BivData)
Curated_BivData <- add_sex(Curated_BivData)
Curated_BivData <- add_subsp(Curated_BivData)
Curated_BivData <- add_species(Curated_BivData)

@@

##change factor cols to numbers (factor -> chr -> number)
Curated_BivData$centromere_ABS_Position <- as.character(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.character(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.character(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.character(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.character(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.character(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.character(Curated_BivData$Foci3)
Curated_BivData$Foci4 <- as.character(Curated_BivData$Foci4)

##Numeric
Curated_BivData$centromere_ABS_Position <- as.numeric(Curated_BivData$centromere_ABS_Position)
Curated_BivData$centromere_PER_Position <- as.numeric(Curated_BivData$centromere_PER_Position)
Curated_BivData$chromosomeLength <- as.numeric(Curated_BivData$chromosomeLength)

Curated_BivData$hand.foci.count <- as.numeric(Curated_BivData$hand.foci.count)

Curated_BivData$Foci1 <- as.numeric(Curated_BivData$Foci1)
Curated_BivData$Foci2 <- as.numeric(Curated_BivData$Foci2)
Curated_BivData$Foci3 <- as.numeric(Curated_BivData$Foci3)
Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
#I think the number of sig figs and rounding changes slightly from the blocks above

#adding more feature calculations (IFD, ect)
Curated_BivData <- add_IFD(Curated_BivData)#fixed a bug
Curated_BivData$IFD1_PER = Curated_BivData$IFD1_ABS / Curated_BivData$chromosomeLength

@@

Curated_BivData$PER_Foci_1 <- Curated_BivData$Foci1 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_2 <- Curated_BivData$Foci2 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_3 <- Curated_BivData$Foci3 / Curated_BivData$chromosomeLength
Curated_BivData$PER_Foci_4 <- Curated_BivData$Foci4 / Curated_BivData$chromosomeLength

@@

#not working
Curated_BivData$dis.cent <- Curated_BivData$Foci1 - Curated_BivData$centromere_ABS_Position
Curated_BivData$dis.cent.PER  <- Curated_BivData$dis.cent / Curated_BivData$chromosomeLength

#add telomere.dist
Curated_BivData$telo_dist <- ifelse(grepl(1, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci1), 
                     ifelse(grepl(2, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength -  Curated_BivData$Foci2), 
                     ifelse(grepl(3, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci3),
                    ifelse(grepl(4, Curated_BivData$hand.foci.count), (Curated_BivData$chromosomeLength - Curated_BivData$Foci4),   0))) )

Curated_BivData$telo_dist_PER <- Curated_BivData$telo_dist / Curated_BivData$chromosomeLength

#sis.co.ten
Curated_BivData <- add_SisCoTen(Curated_BivData)

#add rec.group
#divide musc males into groups
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                                  ifelse(grepl("MSM male", Curated_BivData$category), 1,
                                  ifelse(grepl("SKIVE male", Curated_BivData$category), 1,
                                           0)))

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )
```

>>>>

#BIVDATA MASTER SHEET
new.Curate.FULL <- read.csv("~./MLH1repo/data/BivData/curation/FULL_Curated_BivData.csv")

#found that one of the data sheets is missin bunch of SKIVE female measures
#read in missing SKIVE -- then

#original Curate_bivata + missing SKIVE female
Curated_BivData <-  read.csv2("~./MLH1repo/data/Curated_BivData_5.30.20.csv", sep=" ", header=TRUE)
#reformate this

#83975
#83976
#83009
#80898
#80919
#61084
#60241
#60830

#Clean away all the empty rows
missing.SKIVE.female <- missing.SKIVE.female[!(is.na(missing.SKIVE.female$fileName) | missing.SKIVE.female$fileName==""), ]
#this takes a while to run

#new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file

missing.SKIVE.female <- add_mouse(missing.SKIVE.female)
#make OG copy
new.Curate.FULL.org <- new.Curate.FULL
new.Curate.FULL.org <- add_category(new.Curate.FULL.org)

@@

#missing.SKIVE.female

missing.SKIVE.female=read.xlsx2("~./MLH1repo/data/missing_SKIVE_female.xlsx", header=TRUE, 1)
str(missing.SKIVE.female)

missing.SKIVE.female$fileName <- as.character(missing.SKIVE.female$fileName)

missing.SKIVE.female <- missing.SKIVE.female[!(is.na(missing.SKIVE.female$fileName) | missing.SKIVE.female$fileName==""), ]
#this takes a while to run

#new.Curate.FULL <- add_mouse(new.Curate.FULL)#this function is v.slow for this big file

#SC PASS
missing.SKIVE.female <- missing.SKIVE.female[missing.SKIVE.female$SC.pass == '1',]#
#10480 #9819, #9769

#fix these duplicated ones
#check duplicates
#anyDuplicated(missing.SKIVE.female$Obj.ID)
#0 duplicates for missing SKIVE.f
#remove duplicates
Curated_BivData <- new.Curate.FULL %>% distinct()#from 8933 to 8151

missing.SKIVE.female <- add_mouse(missing.SKIVE.female)
missing.SKIVE.female <- add_category(missing.SKIVE.female)
missing.SKIVE.female <- add_strain(missing.SKIVE.female)
missing.SKIVE.female <- add_sex(missing.SKIVE.female)
missing.SKIVE.female <- add_subsp(missing.SKIVE.female)
missing.SKIVE.female <- add_species(missing.SKIVE.female)

@@

##change factor cols to numbers (factor -> chr -> number)
missing.SKIVE.female$fileName <- as.character(missing.SKIVE.female$fileName)
missing.SKIVE.female$Obj.ID <- as.character(missing.SKIVE.female$Obj.ID)

missing.SKIVE.female$centromere_ABS_Position <- as.character(missing.SKIVE.female$centromere_ABS_Position)
missing.SKIVE.female$centromere_PER_Position <- as.character(missing.SKIVE.female$centromere_PER_Position)
missing.SKIVE.female$chromosomeLength <- as.character(missing.SKIVE.female$chromosomeLength)

missing.SKIVE.female$hand.foci.count <- as.character(missing.SKIVE.female$hand.foci.count)

missing.SKIVE.female$Foci1 <- as.character(missing.SKIVE.female$Foci1)
missing.SKIVE.female$Foci2 <- as.character(missing.SKIVE.female$Foci2)
missing.SKIVE.female$Foci3 <- as.character(missing.SKIVE.female$Foci3)
#missing.SKIVE.female$Foci4 <- as.character(missing.SKIVE.female$Foci4)

##Numeric
missing.SKIVE.female$centromere_ABS_Position <- as.numeric(missing.SKIVE.female$centromere_ABS_Position)
missing.SKIVE.female$centromere_PER_Position <- as.numeric(missing.SKIVE.female$centromere_PER_Position)
missing.SKIVE.female$chromosomeLength <- as.numeric(missing.SKIVE.female$chromosomeLength)

missing.SKIVE.female$hand.foci.count <- as.numeric(missing.SKIVE.female$hand.foci.count)

missing.SKIVE.female$Foci1 <- as.numeric(missing.SKIVE.female$Foci1)
missing.SKIVE.female$Foci2 <- as.numeric(missing.SKIVE.female$Foci2)
missing.SKIVE.female$Foci3 <- as.numeric(missing.SKIVE.female$Foci3)
#Curated_BivData$Foci4 <- as.numeric(Curated_BivData$Foci4)
#I think the number of sig figs and rounding changes slightly from the blocks above

#adding more feature calculations (IFD, ect)
missing.SKIVE.female <- add_IFD(missing.SKIVE.female)#fixed a bug
missing.SKIVE.female$IFD1_PER = missing.SKIVE.female$IFD1_ABS / missing.SKIVE.female$chromosomeLength

@@

missing.SKIVE.female$PER_Foci_1 <- missing.SKIVE.female$Foci1 / missing.SKIVE.female$chromosomeLength
missing.SKIVE.female$PER_Foci_2 <- missing.SKIVE.female$Foci2 / missing.SKIVE.female$chromosomeLength
missing.SKIVE.female$PER_Foci_3 <- missing.SKIVE.female$Foci3 / missing.SKIVE.female$chromosomeLength
missing.SKIVE.female$PER_Foci_4 <- missing.SKIVE.female$Foci4 / missing.SKIVE.female$chromosomeLength

@@

#not working
missing.SKIVE.female$dis.cent <- missing.SKIVE.female$Foci1 - missing.SKIVE.female$centromere_ABS_Position
missing.SKIVE.female$dis.cent.PER  <- missing.SKIVE.female$dis.cent / missing.SKIVE.female$chromosomeLength

#add telomere.dist
missing.SKIVE.female$telo_dist <- ifelse(grepl(1, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci1), 
                     ifelse(grepl(2, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength -  missing.SKIVE.female$Foci2), 
                     ifelse(grepl(3, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci3),
                    ifelse(grepl(4, missing.SKIVE.female$hand.foci.count), (missing.SKIVE.female$chromosomeLength - missing.SKIVE.female$Foci4),   0))) )

missing.SKIVE.female$telo_dist_PER <- missing.SKIVE.female$telo_dist / missing.SKIVE.female$chromosomeLength

#sis.co.ten
missing.SKIVE.female <- add_SisCoTen(missing.SKIVE.female)

#add rec.group
#divide musc males into groups
missing.SKIVE.female$Rec.group <- ifelse(grepl("PWD male", missing.SKIVE.female$category), 1, 
                                  ifelse(grepl("MSM male", missing.SKIVE.female$category), 1,
                                  ifelse(grepl("SKIVE male", missing.SKIVE.female$category), 1,
                                           0)))

#remove bivalent entries without hand foci count values
cleanCurated_BivData <-  Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |
                                      hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 )

#write excel file
#write.xlsx2(missing.SKIVE.female, "Curated_missing.SKIVE.female.xlsx", sheetName = "Sheet1", 
#            col.names = TRUE, row.names = FALSE, append = FALSE)
